Centralized Programming for Distributed Control Planes

### 2016-01-27 paper/compilation.tex
edits to 5

\begin{figure*}
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Topology}{}
  \vspace*{-1\baselineskip}

  \includegraphics[width=.8\columnwidth]{figures/topology}

  \hdr{\large Policy Automata}{}
  \vspace*{1\baselineskip}

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{}
    \state{5}{$5$}{right of=4}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{D}{}
    \transition{2}{3}{C}{}
    \transition{3}{4}{A}{}
    \transition{4}{5}{W}{}
  \end{tikzpicture}

@@

  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Product Graph}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.8\columnwidth]{figures/productgraph}
  \end{minipage}

@@

  \caption{Example Product graph construction.}
  \label{fig:example-compilation}
\end{figure*}

@@

The final stages of compilation consist in translating policies from PGIR to a simple, vendor-neutral abstraction of BGP (ABGP), and then translating from ABGP to actual device configurations.

@@

Having found a total ordering on node preferences in the product graph from the failure safety analysis, the translation to ABGP becomes straightforward. The idea is to encode the state of the automaton using BGP community values. Each router will match based on its peer and a community value corresponding to the state of the product graph, and then update the state before exporting to the neighbors permitted according to the product graph. For example, router $A$ from the example in Figure~\ref{fig:example-compilation} will allow an advertisement from its peer $C$ with a community value for state $(3,2)$ (and deny anything else). If it sees such an advertisement, it will remove the old community value, and add a new community value for state $(4,2)$ before exporting the route to its neighbor $W$.

To ensure preferred paths are always obtained, for each router $r$ in the topology, the compiler will simply set a higher local preference for neighbors of a more-preferred node for $r$ in the product graph. For example, $C$ will prefer an advertisement from $D$ in state $(2,2)$ over an advertisement from $E$ in state $(-,2)$.

Since the compiler is only able to control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it simply translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$ and prefer them equally. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. In practice, we model the unknown AS topology as a special node in the product graph that generates a filter to match any sequence of ASes.

Finally, in our example, the external AS $W$ would need to prefer our internal router $A$ over $B$. In general, it is not possible to control traffic entering the network beyond certain special cases. In this example however, the BGP MED attribute can influence $W$ to prefer $A$ over $B$ since the preference is for a single external peer.  Additionally, the compiler can use the BGP no-export community can ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize various BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).

Figure~\ref{fig:config} shows the full configuration from the compilation example. To make configurations more compact and readable, the \sysname compiler will rewrite configurations in ABGP - for example removing community tagging and matching when there is no ambiguity. Figure~\ref{config-min} shows the minimized router configurations.

@@

As was demonstrated in Example 2 of Section~\ref{sec:motivation}, the use of aggregation in BGP can lead to subtle black-holing traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix going through a router where an aggregate is advertised to go over a single link, even if there are several available in the toplogy. If the single link fails, then a black hole might be introduced. Fortunately, the PGIR captures this information precisely, describing the how information meeting the policy flows over the topology.

We can frame the aggregation problem as a problem of connectivity in the PGIR. More specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures that would disconnect where the prefix originates from where its more specific aggregate is located. The difficulty lies in the fact that the same links in the topology can appear in multiple places in the PGIR. We adopt the following simple strategy to lower bound the number of failures:
\begin{enumerate}
	\item Pick a random path through the product graph from start to end.
	\item Remove all edges in the product graph corresponding to the set of topological edges choosen.
	\item Repeat until no path exists.
\end{enumerate}
Since each path through the PGIR will result in a new edge-disjoint path through the topology (since we have removed all already used edges), and since the number of edge-disjoint paths through the topology is no more than the min-cut. \todo{clarify this}

For example, imagine we use the earlier data center example from Section~\ref{sec:motivation}, with the policy: $p_{g1} \mapsto end(A)$, where $p_{g1}$ falls under the $p_{g^*}$ aggregate. The data center topology and a simplification of the corresponding product graph are shown in Figure~\ref{fig:aggregation-safety}. Since the compiler knows an aggregate will be placed at $X$, and it knows that, for prefix $p_{g1}$ the route will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$ in the product graph. In this case, we could remove the $A$ -- $D$ -- $X$ path first. We would then need to remove any other $A$ -- $D$ or $D$ -- $X$ links from the product graph, though in this case there are none. Next, we could remove the links along the $A$ -- $C$ -- $X$ path, repeating the process. Because afterwards $A$ is disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix $p_{g1}$. This process is then repeated for other aggregation locations (e.g., $Y$).

>>>>

\begin{figure*}
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Topology}{}
  \vspace*{-2\baselineskip}

  \includegraphics[width=.6\columnwidth]{figures/topology}

  \hdr{\large Policy Automata}{}
  \vspace*{-0.5\baselineskip}

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.4cm,minimum size=1cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{}
    \state{5}{$5$}{right of=4}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{D}{}
    \transition{2}{3}{C}{}
    \transition{3}{4}{A}{}
    \transition{4}{5}{W}{}
  \end{tikzpicture}

@@

  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Product Graph IR}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.65\columnwidth]{figures/productgraph}
  \end{minipage}

@@

  \caption{Example Product Graph IR construction.}
  \label{fig:example-compilation}
\end{figure*}

@@

The final stages of compilation translate policies from PGIR to a  vendor-neutral abstraction of BGP (ABGP) and then from ABGP to actual device configurations.

@@

Once we have the total ordering on node preferences in the PGIR from the failure safety analysis, the translation to ABGP is straightforward. The idea is to encode the state of the automaton using BGP community values. Each router will match based on its peer and a community value corresponding to the state of the PGIR, and then update the state before exporting to the neighbors permitted by the PGIR. For example, router $A$ in Figure~\ref{fig:example-compilation} will allow an announcement from $C$ with a community value for state $(3,2)$ (and deny anything else). If it sees such an announcement, it will remove the old community value, and add a new one for state $(4,2)$ before exporting it to $W$.

To ensure preferred paths are always obtained, for each router $r$ in the topology, the compiler sets a higher local preference for neighbors of a more-preferred node for $r$ in the PGIR. For example, $C$ will prefer an advertisement from $D$ in state $(2,2)$ over an advertisement from $E$ in state $(-,2)$.

Since the compiler is only able to control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$ and prefer them equally. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. In practice, we model the unknown AS topology as a special node in the product graph that generates a filter to match any sequence of ASes.

Finally, in our example, the external AS $W$ should prefer our internal router $A$ over $B$. In general, it is not possible to reliably control traffic entering the network beyond certain special cases. In this example, however, assuming our network and$W$ have an agreement to honor MEDs, the BGP MED attribute can influence $W$ to prefer $A$ over $B$. Additionally, the compiler can use the BGP no-export community to ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize various BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).

Figure~\ref{fig:config} shows the full configuration from the compilation example. To make configurations more compact and readable, the \sysname compiler rewrites configurations in ABGP, e.g., removing community tagging and matching when there is no ambiguity. Figure~\ref{config-min} shows minimized router configurations. \todo{figure was removed?}

@@

As demonstrated with Example 2 in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix going through a router where an aggregate is advertised to go over a single link, even if there are several available in the toplogy. If that link fails, a black hole might be introduced. Fortunately, the PGIR captures this information precisely.
%, describing the how information meeting the policy flows over the topology.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures that would disconnect where the prefix originates from where its more specific aggregate is located. The difficulty is that the same links in the topology can appear in multiple places in the PGIR. We adopt the following simple strategy to lower bound the number of failures:
\begin{enumerate}
	\item Pick a random start-to-end path in PGIR.
	\item Remove all edges in the PGIR for the set of topological edges chosen.
	\item Repeat until no path exists.
\end{enumerate}
The number of PGIR paths that we are able to remove is a lower bound on the number of failures. Since each PGIR path results in a new edge-disjoint path through the topology (because of Step 2), and since the number of edge-disjoint paths through the topology is no more than the min-cut. \todo{clarify this}

For example, imagine we use the data center example from \S\ref{sec:motivation}, with the policy: $p_{g1} \mapsto end(A)$, where $p_{g1}$ falls under the $p_{g^*}$ aggregate. Figure~\ref{fig:aggregation-safety} shows the network topology and a simplified PGIR. Since the compiler knows an aggregate will be placed at $X$, and it knows that, the route for $p_{g1}$ will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR, though in this case there are none. Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix $p_{g1}$. This process is repeated for other aggregation locations (e.g., $Y$).

### 2016-01-27 paper/evaluation.tex
adding initial performance numbers of backbone

@@

\subsection{Propane-generated configurations}

Finally, we comment briefly on how the configurations generated by \sysname differ from those generated by operators today. In some cases, \sysname-generated configurations are implemented as one would expect. For example, preferences among neighboring ASes are implemented with a compiler-generated community value to tag incoming routes according to preference, which is then observed at other border routers to influence decisions. In other cases, the \sysname configurations are different, often relying on a different BGP mechanism to achieve the same result. For example, in one instance operators would use the no-export community to prevent routes from leaking beyond a certain tier of the data center, whereas \sysname would selectively import the route only below a certain tier. \sysname could use a similar implementation mechanism in the future as an optimization. In another instance, operators made assumptions about a neighboring AS to cut down on the propogation of route advertisements.

>>>>

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/compilation-times-backbone.png}
\label{fig:compilation-times-dc}
\caption{Data center compilation times.}
\end{figure}

@@

\subsection{Propane vs. operator configurations}

Finally, we comment briefly on how \sysname-generated configurations differ from those generated by operators today. 
%
In some cases, \sysname configurations are implemented as one would expect. For example, preferences among neighboring ASes are implemented with a compiler-generated community value to tag incoming routes according to preference, which is then observed at other border routers to influence decisions. In other cases, the \sysname configurations are different, relying on a different BGP mechanism to achieve the same result. Some key differences that we observed were: 
%
$i)$ operators used the no-export community to prevent routes from leaking beyond a certain tier of the data center, whereas \sysname selectively imported the route only below a certain tier; 
%\sysname could use a similar implementation mechanism in the future as an optimization.
%
$ii)$ operators prevented unneeded propagation of more-specific route advertisements from a neighboring AS based on their out-of-band knowledge about topology, whereas \sysname propagated these advertisements; and
%
$iii)$ operators used a layer of indirection using community groups and re-writing to implement certain policies in a more maintainable way, where \sysname uses flat communities. We are currently investigating if such differences matter to operators (e.g., if they want to read \sysname configurations) and, if necessary, how to reduce them.

### 2016-01-27 paper/introduction.tex

The centralized control planes of SDN, however, are not a panacea.
%
They require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure~\cite{x,y,z}. To address this challenge, many researchers are beginning to explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{x,y,z} and their current programming difficulties.  However, academic
language design and implementation efforts have not kept pace.  For instance, work on
experimental SDN languages
such as Frenetic~\cite{frenetic}, Flowlog~\cite{flowlog}, Vericon~\cite{vericon}, Merlin~\cite{merlin}, NetKAT~\cite{netkat}, and Kinetic~\cite{kinetic}, to
name a few, has not yet shown how to implement fault tolerant
multi-controller systems that support their abstractions efficiently.

In this paper, we ask if it is possible to program distributed control planes using network-wide policies.
Operators should be able to directly express network-wide policies, which are automatically decomposed into individual device configurations. The resulting forwarding behavior, which emerges through device interactions, must be policy-compliant under all possible failures.
If successful, this approach would combine easy programmability of centralized control planes and failure robustness of distributed control planes.
%
More pragmatically, it would help many networks that, for the foreseeable future, will continue to use a distributed control plane, due to the difficulty of migrating to SDN or the inherent scalability and failure-robustness of distributed control.

>>>>

The centralized control planes of SDN, however, are not a panacea.
%
They require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of larger networks and to react sufficiently quickly
to a changing environment. To address this challenge, some researchers are beginning to 
explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.  However, academic
language design and implementation efforts have not kept pace.  For instance, work on
experimental SDN languages
such as Frenetic~\cite{frenetic}, Flowlog~\cite{flowlog}, Vericon~\cite{vericon}, Merlin~\cite{merlin}, NetKAT~\cite{netkat}, and Kinetic~\cite{kinetic}, to
name a few, has not yet shown how to implement fault tolerant
multi-controller systems that support their high-level abstractions efficiently.

In this paper, we ask if it is possible to program 
distributed control planes using network-wide policies.  In particular,
operators should be able to directly express \emph{end-to-end}, 
\emph{network-wide} policies directly; it is the job of a compiler
to decompose such policies into individual device configurations. 
In addition, the resulting forwarding behavior, which emerges through device interactions, must be policy-compliant under all possible failures.
If successful, this approach would combine easy programmability of centralized control planes with the failure robustness and scalability of distributed control planes.
%
More pragmatically, it would help many networks that, for the foreseeable future, will continue to use a distributed control plane, due to the difficulty of migrating to SDN or the inherent scalability and failure-robustness of distributed control.
